---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rror

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/rror)](https://CRAN.R-project.org/package=rror)
<!-- badges: end -->

`rror` is a package to make using classed conditions in R easier. It contains 
utilities for creating, raising, and catching conditions including but not 
limited to messages, warnings, and errors. It is built on top of the base 
condition handling system, and is fully compatible with it.

## Installation

`rror` is not on CRAN yet, but can be installed from GitHub with 

``` r
# install.packages("devtools")
devtools::install_github("alistaire47/rror")
```

## What are conditions?

_Conditions_ are a formalized way of providing feedback to the user and handling 
runtime issues. In most languages, conditions include errors and warnings; in R, 
_messages_ (usually generated with `message()`, equivalent to `warning()` or 
`stop()`) are also conditions, used for generating diagnostic messages for the 
user.

R also has infrastructure to create, raise, and handle conditions of your own 
creation, which may or may not inherit from one of the built-in conditions.

## Why `rror`?

The condition handling system of R is very powerful, but unfortunately most 
users don't take advantage of its full power, because it is not trivial to 
understand, a process further hindered by limited documentation. 

`rror` makes it very simple to leverage more of the power of this system—in 
particular working with classed errors. For instance, if a package maintainer 
is validating arguments and wants to raise an error for bad input, instead of 
generating a `simpleError` from `stop()`, e.g.

```{r}
plus1 <- function(x) {
    if (!is.numeric(x)) {
        stop("`x` must be numeric")
    }
    
    x + 1
}

plus1(2)

tryCatch(
    plus1("a"),
    error = function(e) str(e)
)
```

the maintainer can raise a classed error, e.g.

```{r}
plus2 <- function(x){
    if (!is.numeric(x)) {
        rror::raise(rror::error_condition("`x` must be numeric", class = "value_error"))
    }
    
    x + 2
}

plus2(3)

rror::except(
    plus2("b"),
    value_error = function(e) str(e)    # we can catch specific error classes!
)
```

Common conditions can be predefined and reused.

`rror`'s condition handling function, `except()`, also has some benefits over 
`tryCatch()`. In particular, because the handlers registered in `tryCatch()` are 
exiting handlers, it is not very useful for suppressing repeated warnings or 
messages and continuing:

```{r}
f1 <- function(i) warning("warning ", i)

tryCatch(
    {
        for (i in 1:3) f1(i)
        2 + 2
    },
    warning = function(w) print(w) 
)
```

`except()` handles errors with exiting handlers, but non-fatal conditions use 
local handlers, which lets specific errors be easily suppressed:

```{r}
f2 <- function(i) {
    rror::raise(
        rror::warning_condition(
            paste("warning", i), 
            class = "number_warning"
        )
    )
}

rror::except(
    {
        for (i in 1:3) f2(i)
        2 + 2
    },
    number_warning = function(w) print(w) 
)
```